### 2021-01-23 paper/arm.tex
Egits

\subsection{Down-Grading Acquires ($\xDGA$)}
\label{sec:internal}

\begin{example}
The following execution is allowed by \armeight{}, but disallowed by
\refdef{def:pomsets-ra}.  The coherence order between the writes can be
witnessed by a separate thread, which we have elided.
\begin{gather*}
  \PW{x}{2}\SEMI 
  \PR[\mRA]{x}{r}\SEMI
  \PW{y}{1} \PAR
  \PW{y}{2}\SEMI
  \PW[\mRA]{x}{1}
  \\
  \hbox{\begin{tikzinline}[node distance=1.5em]
      \event{a}{\DW{x}{2}}{}
      \raevent{b}{\DR[\mRA]{x}{2}}{right=of a}
      \event{c}{\DW{y}{1}}{right=of b}
      \event{d}{\DW{y}{2}}{right=2.5em of c}
      \raevent{e}{\DW[\mRA]{x}{1}}{right=of d}
      \rf{a}{b}
      \sync{b}{c}
      \wk{c}{d}
      \sync{d}{e}
      \wk[out=-165,in=-15]{e}{a}
    \end{tikzinline}}
\end{gather*}
Under \EGC{}, this is explained by dropping the order
$\DRP[\mRA]{x}{2}\xsync\DWP{y}{1}$, because $\DRP[\mRA]{x}{2}$ is fulfilled
by a relaxed write in the same thread.
\begin{gather*}
  \hbox{\begin{tikzinline}[node distance=1.5em]
      \event{a}{\DW{x}{2}}{}
      \raevent{b}{\DR[\mRA]{x}{2}}{right=of a}
      \event{c}{\DW{y}{1}}{right=of b}
      \event{d}{\DW{y}{2}}{right=2.5em of c}
      \raevent{e}{\DW[\mRA]{x}{1}}{right=of d}
      \rf{a}{b}
      %\sync{b}{c}
      \wk{c}{d}
      \sync{d}{e}
      \wk[out=-165,in=-15]{e}{a}
    \end{tikzinline}}  
\end{gather*}

More generally, this can be understood as a compiler optimization that
downgrades a read from $\mRA$ to $\mRLX$ when it read can be fulfilled by a
relaxed write in the same thread.
\end{example}

To model such \emph{downgraded acquires}, we use the uninterpreted symbols
$\Dx{\aLoc}$.

@@

\begin{definition}[$\xCO$/$\xRASC$/$\xDGA$]
  \label{def:pomsets-down}
  Update \refdef{def:pomsets-ra} to:
  \begin{enumerate}
  \item[\ref{S4})]
    $\aTr{\bEvs}{\bForm}$ implies $(\Qw{\aLoc}\limplies\aExp{\neq}\aVal)\land \DS{\aLoc}{\amode}\bForm[\aExp/\aLoc]$,
  \item[\ref{S5})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qw{\aLoc}\land \DS{\aLoc}{\amode}\bForm[\aExp/\aLoc]$.
  \item[\ref{L5})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land((\aVal{=}\aReg\lor\aLoc{=}\aReg)\limplies\bForm)$,
  \item[\ref{L6})]
    $\aTr{\dEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land\bForm$.
  \end{enumerate}
\end{definition}

@@

\begin{example}
One might worry that our model is too permissive for $\mSC$ access, but
\armeight{} allows some counterintuitive results for SC access.  In the
following execution we elide the initializing write $\DWP{y}{0}$.
\begin{gather*}
  \IF{\PR{x}{}}\THEN \PW{x}{2} \FI\SEMI
  \PR[\mSC]{x}{r}\SEMI
  \PR[\mSC]{y}{s} \PAR
  \PW[\mSC]{y}{2}\SEMI
  \PW[\mSC]{x}{1}
  \\
  \hbox{\begin{tikzinline}[node distance=1.5em]
      \event{a}{\DR{x}{1}}{}
      \event{b}{\DW{x}{2}}{right=of a}
      \scevent{c}{\DR[\mSC]{x}{2}}{right=of b}
      \scevent{d}{\DR[\mSC]{y}{0}}{right=of c}
      \scevent{e}{\DW[\mSC]{y}{2}}{right=2.5em of d}
      \scevent{f}{\DW[\mSC]{x}{1}}{right=of e}
      \po{a}{b}
      \rf{b}{c}
      %\sync{c}{d}
      \sync{e}{f}
      \wk{d}{e}
      \rf[out=-165,in=-15]{f}{a}
    \end{tikzinline}}
\end{gather*}
Under \EGC{}, this is explained by dropping the order
$\DRP[\mSC]{x}{2}\xsync\DRP{y}{0}$, because $\DRP[\mSC]{x}{2}$ is fulfilled
by a relaxed write in the same thread.
\end{example}

@@

\begin{example}
The following execution is allowed by \armeight{}, but disallowed by
\refdef{def:pomsets-ra}.  
\begin{gather*}
  \begin{gathered}
  x\GETS0\SEMI 
  x\GETS 1\SEMI y^\mRA\GETS1 \PAR r\GETS y\SEMI \IF{r}\THEN s\GETS x\FI
  \\[-.4ex]
  \hbox{\begin{tikzinline}[node distance=1.5em]
      \event{wx0}{\DW{x}{0}}{}
      \event{wx1}{\DW{x}{1}}{right=of wx0}
      \raevent{wy1}{\DW[\mRA]{y}{1}}{right=of wx1}
      \event{ry1}{\DR{y}{1}}{right=2.5em of wy1}
      \event{rx0}{\DR{x}{0}}{right=of ry1}
      \sync{wx1}{wy1}
      \po{ry1}{rx0}
      \rf{wy1}{ry1}
      \wk[in=-15,out=-165]{rx0}{wx1}
      \wk{wx0}{wx1}
    \end{tikzinline}}
\end{gathered}
\end{gather*}
Under \EGC{}, this is explained by dropping the order
$\DRP{y}{1}\xpo\DRP{x}{0}$, because \armeight{} does not include control
dependencies between reads in the locally-ordered-before relation.
\begin{gather*}
  \hbox{\begin{tikzinline}[node distance=1.5em]
      \event{wx0}{\DW{x}{0}}{}
      \event{wx1}{\DW{x}{1}}{right=of wx0}
      \raevent{wy1}{\DWRel{y}{1}}{right=of wx1}
      \event{ry1}{\DR{y}{1}}{right=2.5em of wy1}
      \event{rx0}{\DR{x}{0}}{right=of ry1}
      \sync{wx1}{wy1}
      %\po{ry1}{rx0}
      \rf{wy1}{ry1}
      \wk[in=-15,out=-165]{rx0}{wx1}
      \wk{wx0}{wx1}
    \end{tikzinline}}  
\end{gather*}
\end{example}

@@

We repeat all of the rules for preconditions and transformers,
Only \ref{L5} and \ref{T3} are changed from \refdef{def:pomsets-down}.
We repeat \ref{L4}, although it is unchanged from \refdef{def:pomsets-trans}.

\begin{definition}[$\xCO$/$\xRASC$/$\xDGA$/$\xRRD$]
  \label{def:pomsets-down-rr}
  Update \refdef{def:pomsets-trans} to:
  \begin{enumerate}
  \item[\ref{S3})]
    $\labelingForm(\aEv)$ implies $\QS{\aLoc}{\amode}\land\aExp{=}\aVal$,
  \item[\ref{L3})]
    $\labelingForm(\aEv)$ implies $\QL{\aLoc}{\amode}$.
  \item[\ref{T3})]
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\Q{}][\TRUE/\RW]$ if $\labelingAct_1(\aEv)$ is a write,
    \\
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\Q{}][\FALSE/\RW]$ otherwise.
  \end{enumerate}
  \begin{enumerate}
  \item[\ref{S4})]
    $\aTr{\bEvs}{\bForm}$ implies $(\Qw{\aLoc}\limplies\aExp{\neq}\aVal)\land \DS{\aLoc}{\amode}\bForm[\aExp/\aLoc]$,
  \item[\ref{S5})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qw{\aLoc}\land \DS{\aLoc}{\amode}\bForm[\aExp/\aLoc]$.
  \item[\ref{L4})]
    $\aTr{\bEvs}{\bForm}$ implies $\aVal{=}\aReg\limplies\bForm$, 
  \item[\ref{L5})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land(\RW\limplies(\aVal{=}\aReg\lor\aLoc{=}\aReg)\limplies\bForm)$,
  \item[\ref{L6})]
    $\aTr{\dEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land\bForm$.
  \end{enumerate}  
\end{definition}

@@

%\subsection{Proof Strategy}

>>>>

\subsection{Downgraded Reads ($\xDGR$)}
\label{sec:internal}

\begin{example}
  The following execution is allowed by \armeight{}, but disallowed by
  \refdef{def:pomsets-ra}.  The coherence order between the writes can be
  witnessed by a separate thread, which we have elided.
  \begin{gather*}
    \PW{x}{2}\SEMI 
    \PR[\mRA]{x}{r}\SEMI
    \PW{y}{1} \PAR
    \PW{y}{2}\SEMI
    \PW[\mRA]{x}{1}
    \\
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{a}{\DW{x}{2}}{}
        \raevent{b}{\DR[\mRA]{x}{2}}{right=of a}
        \event{c}{\DW{y}{1}}{right=of b}
        \event{d}{\DW{y}{2}}{right=2.5em of c}
        \raevent{e}{\DW[\mRA]{x}{1}}{right=of d}
        \rf{a}{b}
        \sync{b}{c}
        \wk{c}{d}
        \sync{d}{e}
        \wk[out=-165,in=-15]{e}{a}
      \end{tikzinline}}
  \end{gather*}
  Under \EGC{}, this is explained by dropping the order
  $\DRP[\mRA]{x}{2}\xsync\DWP{y}{1}$, because $\DRP[\mRA]{x}{2}$ is fulfilled
  by a relaxed write in the same thread.
  \begin{gather*}
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{a}{\DW{x}{2}}{}
        \raevent{b}{\DR[\mRA]{x}{2}}{right=of a}
        \event{c}{\DW{y}{1}}{right=of b}
        \event{d}{\DW{y}{2}}{right=2.5em of c}
        \raevent{e}{\DW[\mRA]{x}{1}}{right=of d}
        \rf{a}{b}
        % \sync{b}{c}
        \wk{c}{d}
        \sync{d}{e}
        \wk[out=-165,in=-15]{e}{a}
      \end{tikzinline}}  
  \end{gather*}

  More generally, this can be understood as a compiler optimization that
  downgrades a read from $\mRA$ to $\mRLX$ when it can be fulfilled by a
  relaxed write in the same thread.
\end{example}

To model such \emph{downgraded reads}, we use the uninterpreted symbols
$\Dx{\aLoc}$.

@@

\begin{definition}[$\xCO$/$\xRASC$/$\xDGR$]
  \label{def:pomsets-down}
  Update \refdef{def:pomsets-ra} to:
  \begin{enumerate}
  \item[\ref{S4})]
    $\aTr{\bEvs}{\bForm}$ implies $(\Qw{\aLoc}\limplies\aExp{\neq}\aVal)\land \DS{\aLoc}{\amode}\bForm[\aExp/\aLoc]$,
  \item[\ref{S5})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qw{\aLoc}\land \DS{\aLoc}{\amode}\bForm[\aExp/\aLoc]$.
  \item[\ref{L5})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land((\aVal{=}\aReg\lor\aLoc{=}\aReg)\limplies\bForm)$,
  \item[\ref{L6})]
    $\aTr{\dEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land\bForm$.
  \end{enumerate}
\end{definition}

@@

\begin{example}
  One might worry that our model is too permissive for $\mSC$ access, but
  \armeight{} itself allows some very counterintuitive results for $\mSC$
  access.  In the following execution we elide the initializing write
  $\DWP{y}{0}$.
  \begin{gather*}
    \IF{\PR{x}{}}\THEN \PW{x}{2} \FI\SEMI
    \PR[\mSC]{x}{r}\SEMI
    \PR[\mSC]{y}{s} \PAR
    \PW[\mSC]{y}{2}\SEMI
    \PW[\mSC]{x}{1}
    \\
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{a}{\DR{x}{1}}{}
        \event{b}{\DW{x}{2}}{right=of a}
        \scevent{c}{\DR[\mSC]{x}{2}}{right=of b}
        \scevent{d}{\DR[\mSC]{y}{0}}{right=of c}
        \scevent{e}{\DW[\mSC]{y}{2}}{right=2.5em of d}
        \scevent{f}{\DW[\mSC]{x}{1}}{right=of e}
        \po{a}{b}
        \rf{b}{c}
        % \sync{c}{d}
        \sync{e}{f}
        \wk{d}{e}
        \rf[out=-165,in=-15]{f}{a}
      \end{tikzinline}}
  \end{gather*}
  Under \EGC{}, this is explained by dropping the order
  $\DRP[\mSC]{x}{2}\xsync\DRP{y}{0}$, because $\DRP[\mSC]{x}{2}$ is fulfilled
  by a relaxed write in the same thread.
\end{example}

@@

\begin{example}
  \label{ex:rrd1}
  The following execution is allowed by \armeight{}, but disallowed by
  \refdef{def:pomsets-ra}.  
  \begin{gather*}
    \begin{gathered}
      x\GETS0\SEMI x\GETS 1\SEMI y^\mRA\GETS1
      \PAR
      r\GETS y\SEMI \IF{r}\THEN s\GETS x\FI
      \\[-.4ex]
      \hbox{\begin{tikzinline}[node distance=1.5em]
          \event{wx0}{\DW{x}{0}}{}
          \event{wx1}{\DW{x}{1}}{right=of wx0}
          \raevent{wy1}{\DW[\mRA]{y}{1}}{right=of wx1}
          \event{ry1}{\DR{y}{1}}{right=2.5em of wy1}
          \event{rx0}{\DR{x}{0}}{right=of ry1}
          \sync{wx1}{wy1}
          \po{ry1}{rx0}
          \rf{wy1}{ry1}
          \wk[in=-15,out=-165]{rx0}{wx1}
          \wk{wx0}{wx1}
        \end{tikzinline}}
    \end{gathered}
  \end{gather*}
  Under \EGC{}, this is explained by dropping the order
  $\DRP{y}{1}\xpo\DRP{x}{0}$, because \armeight{} does not include control
  dependencies between reads in the locally-ordered-before relation.
  \begin{gather*}
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{wx0}{\DW{x}{0}}{}
        \event{wx1}{\DW{x}{1}}{right=of wx0}
        \raevent{wy1}{\DWRel{y}{1}}{right=of wx1}
        \event{ry1}{\DR{y}{1}}{right=2.5em of wy1}
        \event{rx0}{\DR{x}{0}}{right=of ry1}
        \sync{wx1}{wy1}
        % \po{ry1}{rx0}
        \rf{wy1}{ry1}
        \wk[in=-15,out=-165]{rx0}{wx1}
        \wk{wx0}{wx1}
      \end{tikzinline}}  
  \end{gather*}
\end{example}

@@

\begin{example}
  Revisiting \refex{ex:rrd1}
  \begin{align*}
    \begin{gathered}
      r\GETS y
      \\[-.4ex]
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{a}{\DR{y}{1}}{}
          \xform{xd}{1{=}r\limplies\bForm}{below=of a}
          \xform{xi}{(y{=}r\lor1{=}r)\limplies\bForm}{below=of xd}
          \xo{a}{xd}
        \end{tikzinline}}
    \end{gathered}
    &&
    \begin{gathered}
      \IF{r}\THEN s\GETS x\FI
      \\[-.4ex]
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{b}{r\mid\DR{x}{0}}{}
        \end{tikzinline}}
    \end{gathered}
  \end{align*}
\end{example}

\refdef{def:pomsets-down-rr} combines all of the features of
\textsection\ref{sec:q}--\ref{sec:arm}:
\begin{definition}[$\xCO$/$\xRASC$/$\xDGR$/$\xRRD$]
  \label{def:pomsets-down-rr}
  Update \refdef{def:pomsets-trans} to:
  \begin{enumerate}
  \item[\ref{S3})]
    $\labelingForm(\aEv)$ implies $\QS{\aLoc}{\amode}\land\aExp{=}\aVal$,
  \item[\ref{L3})]
    $\labelingForm(\aEv)$ implies $\QL{\aLoc}{\amode}$.
  \item[\ref{T3})]
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\Q{}][\TRUE/\RW]$ if $\labelingAct_1(\aEv)$ is a write,
    \\
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\Q{}][\FALSE/\RW]$ otherwise.
  \end{enumerate}
  \begin{enumerate}
  \item[\ref{S4})]
    $\aTr{\bEvs}{\bForm}$ implies $(\Qw{\aLoc}\limplies\aExp{\neq}\aVal)\land \DS{\aLoc}{\amode}\bForm[\aExp/\aLoc]$,
  \item[\ref{S5})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qw{\aLoc}\land \DS{\aLoc}{\amode}\bForm[\aExp/\aLoc]$.
  \item[\ref{L4})]
    $\aTr{\bEvs}{\bForm}$ implies $\aVal{=}\aReg\limplies\bForm$, 
  \item[\ref{L5})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land(\RW\limplies(\aVal{=}\aReg\lor\aLoc{=}\aReg)\limplies\bForm)$,
  \item[\ref{L6})]
    $\aTr{\dEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land\bForm$.
  \end{enumerate}  
\end{definition}
Here, we have repeated all of the rules for preconditions and transformers,
including \ref{L4}, which is unchanged from \refdef{def:pomsets-trans}.
Only \ref{L5} and \ref{T3} are changed from \refdef{def:pomsets-down}.

@@

% \subsection{Proof Strategy}
